% arara: xelatex
% arara: xelatex
% arara: xelatex

% options:
% thesis=B bachelor's thesis
% thesis=M master's thesis
% czech thesis in Czech language
% english thesis in English language
% hidelinks remove colour boxes around hyperlinks

\documentclass[thesis=M,english]{FITthesis}[2012/10/20]

\usepackage[utf8]{inputenc}

\usepackage{graphicx} %graphics files inclusion
% \usepackage{subfig} %subfigures
% \usepackage{amsmath} %advanced maths
% \usepackage{amssymb} %additional math symbols

\usepackage{dirtree} %directory tree visualisation

% ADDED BY ME
%\usepackage{blindtext} % lorem ipsum
%\definecolor{mygray}{gray}{0.6}
%\newcommand{\blind}[1][1]{\textcolor{mygray}{\Blindtext[#1][1]}}

\newcommand{\blind}[1][1]{}

% ALSO ADDED BY ME
\usepackage{todonotes} 
\usepackage{multicol}

\newcommand{\todotext}[1]{\textcolor{red}{\textbf{[[#1]]}}}
\newcommand{\redtext}[1]{\textcolor{red}{[[#1]]}} 


\setlength{\fboxsep}{0.005pt}
\newcommand{\tmpframe}[1]{\fbox{#1}}
\renewcommand{\tmpframe}[1]{#1}
\newcommand{\permanentframe}[1]{\fbox{#1}}
\newcommand{\noframe}[1]{#1}


\let\myCite\cite
\renewcommand\cite{\unskip~\myCite}
\let\myRef\ref
\renewcommand\ref{\unskip~\myRef}

% \usepackage{spverbatim} % verb and verbatim with line breaks - doesn't break what I need

\newcommand\hatx{\^{}}
\newcommand{\hatxxx}{\hatx\hatx\hatx}


% list of acronyms
%\usepackage[acronym,nonumberlist,toc,numberedsection=autolabel,nomain]{glossaries}
\iflanguage{czech}{\renewcommand*{\acronymname}{Seznam pou{\v z}it{\' y}ch zkratek}}{}
%\makeglossaries

\newcommand{\tg}{\mathop{\mathrm{tg}}} %cesky tangens
\newcommand{\cotg}{\mathop{\mathrm{cotg}}} %cesky cotangens

% LB
% \setlength{\parskip}{0em}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\department{Department of theoretical computer science}
\title{Contextual Shell History}
\authorGN{Šimon} %author's given name/names
\authorFN{Let} %author's surname
\authorWithDegrees{Bc. Šimon Let} %author's name with academic degrees
\author{Šimon Let} %author's name without academic degrees
\supervisor{Ing. Lukáš Bařinka}

\acknowledgements{I would like to thank my supervisor Ing. Lukáš Bařinka for his guidance and advice. His insights and expertise shaped this work significantly. 

Next, my thanks goes to my colleagues and friends for their valuable feedback. They helped me spot my personal biases and assumptions rooted in my personal shell workflows. 

I am grateful for the trust of hundreds of people who downloaded this project. I would also like to thank all GitHub users who submitted issues and pull requests to the project.

I want to thank Installfest conference organizers for the opportunity to give a talk and to demo this project on Installfest.  

Finally, I would like to thank my family for their support during writing this thesis and trough my whole studies.
}
\abstractCS{I dnes zůstává příkazová řádka populárním způsob jak ovládat počítač. Historie shellu umožňuje lidem snadno opakovat předchozí příkazy což zvyšuje jejich 
produktivitu.
Způsob jakým lidé používají shell závisí na kontextu jako je třeba současný adresář. V této práci chceme využít dostupný kontext ke zlepšení standartní historie shellu.

%Even nowadays, command line is a popular way to interact with computers. Shell history allows people to reuse previous commands which increases productivity. 
%The way people use shell changes based on context such as current directory. In this work, we intend to use additional context to enhance the standard shell history.

Zkoumáme jak lidé používali shell v minulosti a jak ho používají v současnosti.
Analyzujeme funkce standartní historie a to jak je lidé používají. Identifikujeme interaktivní zpětné vyhledávání jako neefektivní mechanismus standartní historie vhodný pro náhradu.   

%We research how people used shell in the past and how they use it now.
%We analyze standard history features and how people use them. We identify reverse search as inefficient standard feature we need to redesign.

Prozkoumáním existujících nástrojů historie shellu najdeme nástroje, které řeší problémy s interaktivním zpětným vyhledáváním. Nacházíme také kontextové nástroje historie shellu, které ale nepřináší uživateli velké zlepšení.

%By exploring existing history tools we find tools that address the issues of reverse search. We also find contextual history tools that do not bring much value to the user. 

Navrhujeme systém historie shellu, který přináší výhody existujících nekontextových nástrojů historie. Zároveň náš návrh používá kontext k dalšímu zlepšení funkcí historie shellu.

%We design a history system that matches the improvement of state-of-the-art non-contextual history tools; Plus, our design uses context to further enhance the capabilities of the shell history.

Implementujeme podstatnou část našeho návrhu. Naše řešení zaznamenává historii shellu s kontextem. Terminálová aplikace vyhledává v historii a zobrazuje relevantní výsledky na základě současného kontextu. 

%We implement a core parts of the design. It records shell history with context. A fullscreen terminal app searches the history and returns relevant results based on current context.

Na základě dát o používání od našich uživatelů porovnáváme naše řešení s populárním existujícím nástrojem pro vyhledávání v historii -- Hstr. Naše řešení v průměru podává podobný nebo lepší výkon než Hstr.

%Based on usage data from some of our users, we compare our solution with one of the state-of-the-art history tools -- Hstr.
%Our solution on average performs similarly or better than Hstr.

Existuje mnoho situací, kde naše řešení překonává Hstr. Uživatel při vyhledávání potřebuje méně znalosti a musí méně psát. Naše řešení v průměru šetří uživateli více napsaných znaků.

%We found many situations where our solution outperforms Hstr. It requires less knowledge and less typing when searching. Our solution on average saves more typed chars to the user. 

Naše řešení bylo za posledních pět měsíců nainstalováno přes 600 krát. Někteří naši uživatelé dříve používali Hstr a naše aplikace pokrývá všechny jejich předchozí potřeby. Zpětná vazba kterou jsme dostali od komunity je z velké části pozitivní.

%Our solution has been installed over 600 times in last five months. Some of our users previously used Hstr and our application covers all of their previous workflows. We have received overwhelmingly positive feedback from the community.
}

\abstractEN{Even nowadays, command line is a popular way to interact with computers. Shell history allows people to reuse previous commands which increases productivity. 
The way people use shell changes based on context such as current directory. In this work, we intend to use available context to enhance the standard shell history.

% e describe target users for our system and typical workflows people perform. Based on the workflows w
We research how people used shell in the past and how they use it now.
We analyze standard history features and how people use them. We identify reverse search as inefficient standard feature we need to redesign.

By exploring existing history tools we find tools that address the issues of reverse search. We also find out that existing contextual history tools do not bring much value to the user. 
%Useful contextual history tools need to match the improvements provided by non-contextual tools.
 
We design a history system that matches the improvement of state-of-the-art non-contextual history tools; Plus it uses context to further enhance the capabilities of the shell history.

We implement a core parts of the design. It records shell history with context. A fullscreen terminal application searches the history and returns relevant results based on current context.

Based on usage data from some of our users, we compare our solution with one of the state-of-the-art history tools -- Hstr.
Our solution on average performs similarly or better than Hstr.

There are many situations where our solution outperforms Hstr. It requires less knowledge and less typing when searching. Our solution on average saves more typed chars to the user. 

Our solution has been installed over 600 times in last five months. Some of our users previously used Hstr and our application covers all of their previous workflows. We have received overwhelmingly positive feedback from the community.

}
\placeForDeclarationOfAuthenticity{Prague} %where you have signed the declaration
\keywordsCS{Shell, Příkazová řádka, Historie shellu, Nástroje produktivity}
\keywordsEN{Shell, Command line, Shell history, Productivity tools}
\declarationOfAuthenticityOption{1} %select as appropriate, according to the desired license
% \website{https://github.com/curusarn/master-thesis} %optional URL (remove entirely if you have no URL for this thesis) TODO: put the thesis up

\begin{document}

%\newacronym{GUI}{GUI}{Graphical User Interface}
%\newacronym{CLI}{CLI}{Command Line Interface}
%\newacronym{GNU}{GNU}{GNU's Not Unix!}
\begin{introduction}



Classic command line interfaces were mostly replaced by GUIs in consumer software \cite{norman2007ui}. However, among IT professionals and enthusiasts, CLIs and shells are a popular way to interact with computers.\footnote{CLIs and shells are active topics on websites that are popular among developers. Specifically, GitHub Topics, Stack Overflow, and Super User all register notable activity under topics/tags about CLIs, Bash, and Zsh.} 
%CLIs are generally not replaceable by GUIs. GUIs do not work well when the number of available actions is high \cite{norman2007ui}. 

Shell history enables shell users to reuse previously entered command line submissions.
There are many people who care about shell history tools.\footnote{We have sparked a wave of reactions by asking people about what shell history tools they are using.\cite{twitter-thread}}

Shells like Bash and Zsh were initially released around year 1990. History features in these shells have not seen many improvements since beginning of the millennium. In this work, we analyze the standard history mechanisms to find how they could be improved.

People use shell differently in different situations. For example, some directories contain projects and people might use different commands based on which project they are working on. We think that we could use directories and more additional information to improve the capabilities of shell history. 

We focus on Unix-based operating systems because more than half of all developers use them\footnote{According to \cite{stackoverflow2019devsurvey}, 25.6\% and 26.8\% of developers use Linux and MacOS respectively.}. 
Default shell on GNU/Linux is Bash \cite{ramey1994gnubash} and default shell on MacOS is Zsh\footnote{According to \cite{apple2019zsh}, default shell in MacOS is Zsh since October 2019.}. Because of that, we focus on these two popular shells.

We research how people used shell in the past and compare it with how they use it now.
Then, we look at standard history features and what issues people encounter when using them. We introduce target users and formulate typical shell history workflows. We explore existing history tools and how they improve standard shell history. Then, we design a new contextual history system based on our analysis. We implement core parts of the designed system. Finally, we test the system using data we collect from our users. Based on that, we draw conclusion about this work.

%\redtext{We analyze: how people use history and what common history-related issues they encounter. We model target users and workflows. We look at existing history tools.}


%About half of all developers have been working with Linux during year 2018. \cite{stackoverflow2019devsurvey} 

%The standard shell history contains records with basic information about previously executed command lines. Each record only contains following details: the command line, timedate of the execution, and the duration of the execution\footnote{Duration is only available in Zsh and only when it is enabled.\cite{zshdocs}}. 

%Additionally, it is quite common to use shell configuration\footnote{\todotext{e.g. Bash: histocontrol=ignoredups\cite{bashman}; zsh: HIST\_SAVE\_NO\_DUPS\cite{zshdocs}}} to prune duplicates from the history. This removes more information from history because the original command line sequences are lost.

%As we can see the standard shell history contains only very basic information about executed command lines.
%There is much more contextual data available at the time of command line execution e.g. present working directory, exit status, and previously executed commands. In this work, we explore the possibilities of using the contextual data about executed commands to increase the usefulness of shell history.



% BAD: This limits the potential of standard shell history utilities because shell usage is contextual. is not generally used to run ad-hoc commands but 

% MORE PRACTICAL: When using shells you likely do not run individual commands completely independently of each other. But instead run commands as parts of different workflows that 

%\todo{describe the contents of this work / layout - redo because moved chapters in analysis} We start by analysing shell history features to understand advantages and disadvantages of standard shell history. We research existing history tools to see what ideas were already explored, which features work and which do not. 

%After that, we focus on people. We start broad and research how people use computer tools in general. Then we move on to examining how people use shell and shell history. We also look at public discourse related to history tools and engage in online discussions to find out more about configurations and tools people use.

%Next, we focus on the contextual information; We cover what information is available to be saved at the time of command line execution and we discuss how it could be used to enhance shell history. 

%At this point we move on to design(ing) ...
%\todotext{TODO: finish this intro}



\end{introduction}


% \begin{verbatim}
% ____5___10___15___20___25___30___35___40___45___50___55___60__64
% \end{verbatim}
% The page is 64 verbatim characters wide


\input{main_analysis}

% DESIGN NOTES

\todotext{XXX: Guidelines to mention/cite in the text}

\todotext{Response times - usability}

https://www.nngroup.com/articles/response-times-3-important-limits/

0.1s - “system is reacting instantaneously” -$>$ prompt - recording

1.0s - “user's flow of thought to stay uninterrupted” -$>$ cli app prototype maximum load time

10s - ...
\todotext{Visibility of system status}
X1: Visibility of system status
system should communicate its state -> highlights, all info displayed
challenge with more underlying info

\todotext{X4: Consistency and standards}

\todotext{Discoverability}

bad in shell

related: recognition rather than recall, consistency and standards
important for user experience, adoption, learning curve
generally hard in shell (recall-based)
-$>$ don’t break expectations (conform to standards)
e.g. new keybinds are not a great idea
make use of habits people could already have built while using std shell history features
e.g. people use arrows to get recent commands and ctrl-R to get arbitrary commands

\todotext{Minimal, simple}

guideline from NNgroup

“Overloading a reuse facility with complex functionality would not make it better.” greenberg


\input{main_design}

\input{main_implementation}

\input{main_testing}

\begin{conclusion}

\redtext{Intro the problem, explain why we solved it and THEN describe how}

We have researched how people used shells in the past and how they use it now.
%??? We have collected and analysed shell history or several people.
The average number potential retrieved characters increased four times since the original study was conducted (around 1993). We have identified that sequential properties of shell history are connected with real specific workflows (inconclusive?). These findings reinforced the idea that it makes sense to design and develop new contextual history tools. 
\redtext{................. Everything above this is a draft \hatxxx ..................}

We have explored shell history features available in current standard shells and pointed out their disadvantages. Market research was conducted to find out what issues people encounter and how they use existing shell history features. 

Based on collected usage data and market research we described the target users of shell history tools. For these target users we have formulated typical workflows that accomplish different tasks using shell history. We have provided a detailed explanation of these workflows pointing out issues with standard shell history mechanisms. As the main finding, we have identified reverse search as very inefficient history feature that is worth redesigning.

We have explored what popular history tools are available. We have found state-of-the-art non-contextual history tools that provide major improvement over standard history by replacing reverse search. These tools address the specific issues of reverse search we identified. There are some contextual history tools that are based on appealing ideas; These tools however, do not match the improvements provided by the state-of-the-art non-contextual history tools. We have described how available contextual information relates to ways people use shell. \redtext{TODO: shorten analysis above as necessary}


% design

Based on our previous analysis, we have designed a history system. The design is a result of iterative process that involved feedback from users. Our design matches the improvements of the state-of-the-art non-contextual history tools. Furthermore, it uses context to enhance the capabilities of the shell history. The design was tested to make sure that it fulfills all requirements. 

% implementation

We have implemented a significant portion of the design. The working solution records shell history with context and usage. It features a fullscreen terminal application that searches shell history and shows relevant results based on current context. Without any search query the application offers history entries from current context. 

% evaluation/testing


We have evaluated the performance of the search application and compared it with both state-of-the-art history tool and standard builtin history search mechanism. 
We have suggested metrics and used them to compare our solution with state-of-the-art history searching tool -- Hstr \cite{toolshstr}.
The average performance of our history searching application is either similar or better than Hstr. 

%We have compared the tools using real life scenarios

We have found real life situations where our solution provides significant advantage over both standard reverse search and Hstr.
Our solution enables the user to search with significantly less knowledge and typing. Furthermore, our search application on average saves more typed characters to the user. 
We have concluded that contextual search is a great default because it offers new possibilities without degrading the average performance.


As a result of feedback of our users, we have added several features to the search application. Notably, users can switch between default contextual search and non-contextual search. This is great for specific situations where using context does not bring value.

Since the release of the first prototype the search application has been installed over 600 times. The project has over 250 stars on GitHub. The feedback we have received is overwhelmingly positive. We have also presented one of the first versions of the application on Installfest conference. \cite{installfest-talk}\cite{installfest} 

Right now, we want to encourage you, the reader, to try out the project for yourself. You can find installation instructions on the GitHub page of the project: \url{https://github.com/curusarn/resh} %\cite{resh-github-homepage}


\paragraph{Future work}

Our work implements the core functionality of the designed history system. However, there are parts of the design that should be implemented because they provide considerable value to the users.

One part of the design we want to point out is synchronization.
Implementing the synchronization connectors for the daemon would enable history searching across devices. It would also make the stored history replicated which could prevent potential data loss.

Outside of the things we have included in the design, we suggest an improvement for our search application based on existing research. According to \cite{greenberg1993computer}, history entries that were already retrieved should be easier retrieve in the future. The scoring function should be modified to include this extra factor. 

Additionally, we propose a topic for future research. Using machine learning techniques and context to provide Autosuggestions tailored to the user is a promising topic for a thesis.   

More research should be done to explore relationships between history entries. Exploring how terminal sessions and sequences of history entries relate to workflows of people could bring useful insights. These insights could be used to improve history searching tools.

\end{conclusion}

\bibliographystyle{iso690.bst}
\bibliography{ref}

\appendix

%\printglossaries

\chapter{Contents of attached medium}\label{app:SDcontent}

\todotext{TODO: Visualise the contents of enclosed media. Use of dirtree is recommended. Note that directories src and text with appropriate contents are mandatory.}


\begin{figure}
	\dirtree{%
		.1 readme.txt\DTcomment{the file with CD contents description}.
		.1 data\DTcomment{the data files directory}.
		.2 graphs\DTcomment{the directory of graphs of experiments}.
		.3 *.eps\DTcomment{the B/W graphs}.
		.3 *.png\DTcomment{the color graphs}.
		.3 *.dat\DTcomment{the graphs data files}.
		.1 exe\DTcomment{the directory with executable WBDCM program}.
		.2 wbdcm\DTcomment{the WBDCM program executable (UNIX)}.
		.2 wbdcm.exe\DTcomment{the WBDCM program executable (Windows)}.
		.1 src\DTcomment{the directory of source codes}.
		.2 wbdcm\DTcomment{the directory of WBDCM program}.
		.3 Makefile\DTcomment{the makefile of WBDCM program (UNIX)}.
		.2 thesis\DTcomment{the directory of \LaTeX{} source codes of the thesis}.
		.3 figures\DTcomment{the thesis figures directory}.
		.3 *.tex\DTcomment{the \LaTeX{} source code files of the thesis}.
		.1 text\DTcomment{the thesis text directory}.
		.2 thesis.pdf\DTcomment{the Diploma thesis in PDF format}.
		.2 thesis.ps\DTcomment{the Diploma thesis in PS format}.
	}
\end{figure}


\end{document}
